Compositional Certified Resource Bounds\ifshort{\vspace{-.15cm}}{}

### 2014-10-23 paper/main.tex
work on the intro

Static analysis of quantitative properties of imperative
programs is an active research area and recent years
have seen many innovations.  Notable tools that
have been developed include SPEED~\cite{GulwaniMC09},
KoAT~\cite{BrockschmidtEFFG14}, PUBS~\cite{AlbertAGPZ12},
Rank~\cite{AliasDFG10}, and LOOPUS~\cite{SinnZV14}.

These tools can derive impressive results on realistic
system software.  However, there are still theoritical
shortcomings that remain open problems to this day:

\begin{itemize}
\item The computed bounds are often
  \emph{non-compositional} local bounds (for a single
  (nested) loop) that are difficult to combine to
  global whole program bounds.
\itemskipIn
\item When an automatic tool fails on a loop of the
  program, no user interaction is possible to derive
  a bound on the full program.
\itemskipIn
\item Existing tools work on a highly abstracted
  version of the program, thus, are unable to provide
  a simple syntactic proof of the correctness of their
  bound.
\itemskipIn
\item Maybe we can mention resource freeing                    XXX
\end{itemize}

This paper presents a new framework to derive worst-case
resource bounds for C programs.  Our choice of C is
mainly motivated by its strong presence in system and
embedded software.  These applications often have to meet
strict time and memory constraints.  To ensure that these
requirements are met, profiling is not enough because non
exhaustive.  In contrast, static analysis tools provide
stronger and sound guarantees.

@@

The cost model and the quantitative Hoare logic
presented are based on a recent work~\cite{veristack14}
that formally verifies bounds on the stack usage of
C programs.  We generalize this work by allowing to
reason on any resource.  The automation, which is the
keystone of this paper, is based on amortized analysis
techniques~\cite{jan-something}.  It makes the analysis
of large programs tractable by automatically generating
resource bounds in the previous logic on a large class
of programs.

@@

>>>>

Static analysis of quantitative properties of imperative
programs is an active research area and recent years
have seen many innovations.  Notable tools that
have been developed include SPEED~\cite{GulwaniMC09},
KoAT~\cite{BrockschmidtEFFG14}, PUBS~\cite{AlbertAGPZ12},
Rank~\cite{AliasDFG10}, and LOOPUS~\cite{SinnZV14}.
These tools can derive impressive results on realistic
system software.  However, there are still theoritical
shortcomings that remain open problems to this day:

\begin{itemize}
\item The computed bounds are often
  \emph{non-compositional} local bounds (for a single
  (nested) loop) that are difficult to combine to
  global whole program bounds.
\itemskipIn
\item When an automatic tool fails on a loop of the
  program, no user interaction is possible to derive
  a bound on the full program.
\itemskipIn
\item Existing tools work on a highly abstracted
  version of the program, thus, are unable to provide
  a simple syntactic proof of the correctness of their
  bound.
\itemskipIn
\item Resources that can be freed like memory cannot
  be handled with the usual techniques.
\end{itemize}

This paper presents a new framework to derive worst-case
resource bounds for C programs.  Our choice of C is
mainly motivated by its strong presence in system and
embedded software.  These applications often have to meet
strict time and memory constraints.

XXX [this has nothing to do here]    To ensure that these
requirements are met, profiling is not enough because non
exhaustive.  In contrast, static analysis tools provide
stronger and sound guarantees.

@@

The cost model and the quantitative Hoare logic
presented are based on a recent work~\cite{veristack14}
that formally verifies bounds on the stack usage of
C programs.  We generalize this work by allowing
to reason on any resource (e.g. evaluation steps,
memory, or stack usage).  The automation, which is the
keystone of this paper, is based on amortized analysis
techniques~\cite{jan-something}.  It makes the analysis
of large programs tractable by automatically generating
resource bounds that are certified in the previous logic
on a large class of programs.

@@

The main innovation that makes amortized analysis
work on imperative languages is to track multivariate
quantites instead of simple functions of program
variables.  This leads to precise bounds expressed as
functions of sizes $|[x,y]| = \max(0,y-x)$ of intervals.
